Ignition of Fuel Beds by Firebrands

### 2022-04-13 litreview.tex
Added to the introduction and the results/discussion of manuscript 3(species)

    Another approach that has been used to predict the onset of ignition in fuels is defining the critical mass flux for ignition. The critical mass flux model asserts that if energy is transferred into a fuel in at such a rate or quantity that the mass flux of pyrolysis products release are above the "critical" value ignition will occur~\cite{Nelson1995}. The critical mass flux approach has been primarily considered for the ignition of solid materials such as PMMA~\cite{Rich2007} and wood slabs~\cite{Yashwanth2015, McAllister2013}. In these studies the critical mass flux was well correlated with ignition however, a general model that predicts ignition across materials and experiments is lacking. One shortcoming is the sensitivity to environmental conditions and material properties. In work conducted by McAllister for wooden plates sensitivities to thermal boundary conditions were anticipated to be magnified in the consideration of fuel beds made of particles. Differences in pyrolysis between the materials was also of concern as oxygen availability and boundary layer effects of the particle fuel beds was postulated to have a significant effect~\cite{McAllister2013}.  Numerical studies considering the critical mass flux for a piloted ignition reported that the mass flux at ignition was sensitive to the Damkolher number above the fuel bed~\cite{Dai2013} giving further support of the sensitivity to boundary conditions (i.e., wind near the ignition source). While further study and development of the critical mass flux model may be adjusted to address these sensitivities it also has other shortcomings.

>>>>

    Another approach that has been used to predict the onset of ignition in fuels is defining the critical mass flux for ignition. The critical mass flux model asserts that if energy is transferred into a fuel in at such a rate or quantity that the mass flux of pyrolysis products release are above the "critical" value ignition will occur~\cite{Nelson1995}. The critical mass flux approach has been primarily considered for the ignition of solid materials such as PMMA~\cite{Rich2007} and wood slabs~\cite{Yashwanth2015, McAllister2013}. In these studies the critical mass flux was well correlated with ignition however, a general model that predicts ignition across materials and experiments is lacking. One shortcoming is the sensitivity to environmental conditions and material properties. In work conducted by McAllister for wooden plates sensitivities to thermal boundary conditions were anticipated to be magnified in the consideration of fuel beds made of particles. Differences in pyrolysis between the materials was also of concern as oxygen availability and boundary layer effects of the particle fuel beds was postulated to have a significant effect~\cite{McAllister2013}.  Numerical studies considering the critical mass flux for a piloted ignition reported that the mass flux at ignition was sensitive to the Damkolher number above the fuel bed~\cite{Dai2013} giving further support of the sensitivity to boundary conditions (i.e., wind near the ignition source). While further study and development of the critical mass flux could potentially be expanded to address its shortcomings other models have been developed that are more promising.

    Proposed ignition models that provide greater accuracy than the hot spot and critical mass flux models do so at the expense of complexity and computational cost. One of the most recent, and most promising, tools for ignition modeling has been developed by Lautenberger and is a customized computational fluid dynamic model.

    Unfortunately, this model and models like it are too time consuming given current computational resources to be applied in across the vast number of conditions and configurations of possible ignition threats in wildland urban interface regions around the globe. Thus, a useable and accurate model will likely land somewhere between the critical mass flux and full CFD model with respect to complexity. Currently, the understanding of which effects are the most influential to ignition is not complete enough to develop such a model. Bridging this knowledge gap by identifying an quantifying which parameters are the most influential to an ignition or non-ignition outcome is the overall goal of this work.

### 2022-04-13 manuscript3.tex
Added to the introduction and the results/discussion of manuscript 3(species)

@@

    The 50\% ignition probability results for the 5.8\si{\meter\per\second} wind speed tests are shown in Table~\ref{tab:composition50temp} along with the results for the 0.1\si{\meter\per\second} tests. From these results there are three observations of note. First, the increase of wind lowered the threshold for ignition probability for the Douglas-fir wood and Pine wood by approximately 30\%. Second, in contrast to the Douglas-fir and Pine wood results, an increase in wind resulted in a decrease in a higher threshold for ignition.  Third, the materials that were not observed to ignite in the quiescent tests were also not observed to ignite in the presence of wind.

    \begin{table*}[hpbt]
        \caption{Heater temperature required for 50\% probability of ignition for the materials and conditions tested.}
        \centering
        \begin{tabular}{crrr}
            % \rowcolor{gray!50}
            Material & 0.1 \si{\meter\per\second}(\si{\celsius}) & 5.8 \si{\meter\per\second} (\si{\celsius}) & $\Delta$T (\si{\celsius})\\
            \hline
            Douglas-fir Wood & 574  & 391 & -185 \\
            Pine Wood        & 634  & 430 & -204\\
            Oak Wood         & 663  & \textgreater750 & 87\\
            Wheat Straw      & \textgreater750 & \textgreater750 & -\\
            Pine Bark        & \textgreater750 & \textgreater750 & -\\
            Douglas-Fir Bark & \textgreater750 & \textgreater750 & -
        \end{tabular}
        \label{tab:composition50temp}
    \end{table*}

>>>>

        \caption{Average bulk density of materials for each material and wind speed.}
        \begin{tabular}{crr}
            Material & U\textsubscript{bulk} = 0.1\si{\meter\per\second} (\si{\kilogram\per\cubic\meter})& U\textsubscript{bulk} = 5.8\si{\meter\per\second} (\si{\kilogram\per\cubic\meter})\\
            Oak Wood         & 344 & 428 \\
            Douglas-fir Wood & 99  & 74 \\
            Wheat Straw      & 85  & 85 \\
            Pine Wood        & 114 & 115 \\
            Pine Bark        & 215 & 172 \\
            Douglas-Fir Bark & 131 & 146  

@@

    The 50\% ignition probability results for the 5.8\si{\meter\per\second} wind speed tests are shown in Table~\ref{tab:composition50temp} along with the results for the 0.1\si{\meter\per\second} tests. From these results there are three observations of note. First, the increase of wind lowered the threshold for ignition probability for the Douglas-fir wood and Pine wood by approximately 30\%. Second, in contrast to the Douglas-fir and Pine wood results, an increase in wind resulted in a decrease in a higher threshold for ignition.  Third, the materials that were not observed to ignite in the quiescent tests were also not observed to ignite in the presence of wind. While the trends for douglas-fir wood and pine wood match those presented in Chapter~\ref{part:manuscript2} and trends reported in other studies of pine needle ignition~\cite{Wang2017} and eucalyptus bark~\cite{Ellis2011, Ellis2015}. The decreasing ignition probability of the oak wood contradicts those observed in the aforementioned studies and this study but is not unprecedented. Ignition studies of various litter layers types ignited by various single firebrands observed a sensitivity to ember location where embers that landed on top of a fuel bed in windy conditions were less likely to cause ignition than those embedded where ignition was enhanced~\cite{Plucinski2008}. This trend was attributed to increased heat loss due to wind. In this study, however, the location of the cartridge heater was consistent across all materials, the energy is not a function of the wind speed, and the flow field around the embers is consistent.  
    \begin{table*}[hpbt]
        \caption{Heater temperature required for 50\% probability of ignition for the materials and conditions tested.}
        \centering
        \begin{tabular}{crrr}
            % \rowcolor{gray!50}
            Material & 0.1 \si{\meter\per\second}(\si{\celsius}) & 5.8 \si{\meter\per\second} (\si{\celsius}) & $\Delta$T (\si{\celsius})\\
            \hline
            Douglas-fir Wood & 574  & 391 & -185 \\
            Pine Wood        & 634  & 430 & -204\\
            Oak Wood         & 663  & \textgreater750 & 87\\
            Wheat Straw      & \textgreater750 & \textgreater750 & -\\
            Pine Bark        & \textgreater750 & \textgreater750 & -\\
            Douglas-Fir Bark & \textgreater750 & \textgreater750 & -
        \end{tabular}
        \label{tab:composition50temp}
    \end{table*}

    Comparing the ignition trends of oak wood to the non ignition of pine bark and douglas-fir bark it would seem that a potential cause of decreased ignition is correlated to the material properties. Density particle size chemistry
